Covid-19 Fields Survey Data Protocol

### 2023-01-25 protocol/Background_and_methods_Jan_24.qmd
Working on model and results

#make variables of interest factors
clean_data<-clean_data %<>%
  mutate_at(c("Income","Race","Age_group","location","Geographic_area","Geographic_area_title"),factor)
# recode the response of interest for categorical analysis
clean_data<-clean_data %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

@@

clean_data <-clean_data %>%
  filter(Health_Region !="North East", Health_Region !="North West")

@@

clean_data %>%
  tbl_summary(
    include=c(
    Health_Region,)
  ) %>% bold_labels()

@@

Health_Regions<-data.frame(Health_Region=c(
  #"North West",
  #"North East",
  "West",
  "East",
  "Central",
  "Toronto"),
  Freq=c(#232299,
    #557000,
    4000000,
    3700000,
    5000000,
    1400000
  )
)

@@

| Variable             | Values                                                                                                                          |
|:----------------|-------------------------------------------------------|
| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
                                                                                                                      |

@@

The cleaning process also included removing outliers that were identified during the preliminary analyses. Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000. After cleaning the dataset contained 5,247 entries (unique respondents).

@@

For each survey participant certain data was automatically captured. This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset. Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset. Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN. Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset. Finally, we assigned the corresponding Health Regio to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).

@@

## Statistical models

Because of the binary outcome of the survey answer of our interest (vaccination status as "yes" or "no") we used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location. 
Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021;@cnat2022a]. At the same time in the case of Canada, others have indicated an ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].

Therefore, we built different logistic regression models that accounted for different levels of interaction between the available socio-economic factors obtained from the Fields Covid survey. The different models are described below.

### Model building and selection

The first model (`model1`) examined the interaction of income and Health Region as it has been shown that there has been an increase in income inequality within the provinces of Canada over time [@marchand2020], and the relatively recent implementation of the Health Regions motivated us to explore if income disparity may be an important factor to consider within each of the Health Regions.

The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking. The model appears in @eq-model1.

$$
\begin{aligned}
\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} +\\ \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income} + \beta_5\textrm{(Health Region} \times \textrm{Income)}
\end{aligned}
$$ {#eq-model1}

Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine.

The second model (`model2`) incorporated an interaction effect between age group and income to examine if a significant interaction occured between these factors as well. The model appears in @eq-model2.

$$
\begin{aligned}
\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \\\beta_4 \textrm{ Income}+ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)}
\end{aligned}
$$ {#eq-model2}

A third model (`model3`) incorporated an interaction effect between race and income, and appears in @eq-model3.

@@

Although it could have been interesting to examine the interaction of Health Region and Race, or 3-way interactions between socio-economic factors (e.g., age,income,race), these models had singularity issues. Next, we compared the AIC of the three models indicated above. The AICs are very similar and appear in @tbl-AIC.

@@

## the first model is a simple logistic regression with fixed effects only svyglm uses the glm
## function to calculate, so it does not allow for random effects.

model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
              design=a1_rake,
              family = quasibinomial(), 
              control= list(maxit=25))

model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
              design=a1_rake,
              family = quasibinomial(), 
              control= list(maxit=25))

@@

#summary(model1.test)
#summary(model1.hr)

model_comp<-as_tibble(AIC(model1,model2,model3))%>%
  select(-c(eff.p,deltabar))%>%
  add_column(model=c("model1","model2","model3"))

model_comp%>%
  kbl(booktabs=T,
      col.names = c("AIC","model"))

## helps visualize the model output

@@

data_summary %>%
  modify_header(label = "**Variable**") %>%
  as_flex_table()

@@

## Model Comparisons

The output from the three models is too long to give a summarized view. I am still trying to get the important parts out but in the meantime here are the summary findings:

- **Model 1**
  Significant values on:
- Age: None
- Race: East Asian/Pacific Islander, South Asian, White Caucasian
- Income: 60000-89000 CAD, Over 90000
- Interaction of Income and Health Region: Toronto and 60000-899999, North West and over_90000, North West  and under_15000

- **Model 2**
    Significant values on:
- Age: None
- Race: East Asian/Pacific Islander, South Asian, White Caucasian
- Health Region: East
- Income: 60000-89000 CAD, Over 90000
- Interaction of Income and Health Region: East and 60000_89999, Toronto and 60000-899999, North West and over_90000, North West  and under_15000
- Interaction of Age and income: A couple of low values, but none below 0.05

including here the results of the third model, which I believe is the most appropriate as it has a similar AIC to the other ones, and it is able to provide interaction information on three different instances.

- **Model 3**
    Significant values on:
- Age: None
- Race: Indigenous
- Health Region: East, Toronto
- Income: 60000-89000 CAD
- Interaction of Income and Health Region: East and 40000_59999, East and 60000_89999, Toronto and 60000-899999, East and  over_90000, North West and over_90000, North West  and under_15000
- Interaction of Age and income: A couple of low values, but none below 0.05
- Interaction of Race and Income: indigenous and 25000_39999, indigenous and over_90000, mixed and over_90000, white_caucasian and over_90000. A visual representation of the estimates of the models can be see below, where @fig-fixed-effects shows the odds-ratio, 95\% confidence interval for the single fixed effects of the three models. @fig-interaction-effects shows the odds ratio and 95\% confidence interval for the interaction effects of models 1 and 2, and @fig-interaction-effects-mod3 shows the interaction effects for model 3.

```{r, model-fixed-eff}
#| message: false
#| label: fig-fixed-effects
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-cap: "Fixed effects of all models"
#| fig-height: 12

 # models<-list("First Model"=model1,
 #              "Second Model"=model2,
 #              "Third Model"=model3)

# cm<-c('Incomeunder_15000'='Under CAD 15000',
#       'Income25000_39999'='CAD 25000-39999',
#       'Income40000_59999'='CAD 40000-59999',
#       'Income60000_89999'='CAD 60000-89999',
#       'Incomeover_90000'='Over CAD 90000',
#       'Age_group25_24'='Age 25-34',
#       'Age_group35_44'='Age 35-44',
#       'Age_group45-54'='Age 45-54',
#       'Age_group55_64'='Age 55-64',
#       'Age_group65_and_over'='Age 65 and over',
#       'Raceblack'='Black',
#       'Raceeast_asian_pacific_islander'='East Asian/Pacific Islander',
#       'Raceindigenous'="Indigenous",
#       'Racelatin_american'='Latin American',
#       'Racemixed'='mixed',
#       'Raceother'='Other',
#       'Racesouth_asian'='South Asian',
#       'Racewhite_caucasian'='White Caucasian',
#       'Health_RegionNorth East'='North East',
#       'Health_RegionNorth West'='North West',
#       'Health_RegionToronto'='Toronto',
#       'Health_RegionWest'='West',
#       
#       )

a<-modelplot(model1,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")+xlab("")

b<-modelplot(model2,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
  scale_color_manual(values = c("black", "red"))+ggtitle("Second Model")+xlab("")

c<-modelplot(model3,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
  scale_color_manual(values = c("black", "red"))+ggtitle("Third Model")

a+b+c+plot_layout(guides = 'collect',ncol=1)+plot_annotation(title="Fixed effects of all models")& theme(legend.position = 'bottom')

#  modelsummary(models,exponentiate = TRUE, stars = T, statistic = c("conf.int",
#                            "p = {p.value}"))
# 
# 
#  modelsummary(model1,exponentiate = TRUE, stars = T, statistic ="conf.int")
# 
#  model_tbl1<-model1 %>%
#    tbl_regression(exponentiate = TRUE) %>%
#    bold_labels()%>%
#    modify_header(label = "**Variable**")
# 
#   test<-summary(model1)
# 
#  model_tbl1<-model1 %>%
#    tbl_regression(exponentiate = TRUE) %>%
#    select(variable, estimate, p.value) %>%
#    filter(p.value <= 0.05)%>%
#    kbl()%>%
#    as_kable_extra() %>%
#    kableExtra::kable_styling(
#      font_size = 8
#    )
# 
# 
# 
# # # 
# # model_tbl2<-model2 %>%
# #   tbl_regression(exponentiate = TRUE) %>%
# #   bold_labels()%>%
# #   modify_header(label = "**Variable**")%>%
# #   as_kable_extra() %>%
# #   kableExtra::kable_styling(
# #     font_size = 8
# #   )
# # 
# model_tbl3<-model3 %>%
#   tbl_regression(exponentiate = TRUE) %>%
#   bold_labels()%>%
#   modify_header(label = "**Variable**")%>%
#   as_kable_extra() %>%
#   kableExtra::kable_styling(
#     font_size = 10
#   )
# 
# # model_tbl3
# # 
# # model_tbl3
# # 
# # tbl_merge_model<-
# #   tbl_merge(
# #     tbls = list(model_tbl1,model_tbl2,model_tbl3),
# #     tab_spanner = c("**Model 1**","**Model 2**", "**Model 3**")
# #   )%>%
# #   as_flex_table()
# # 
# # tbl_merge_model
# # 
# 
#  modelsummary(model3,exponentiate = TRUE, 
#               stars = T, 
#               estimate = "{estimate} [{conf.low}, {conf.high}]",
#               #statistic = NULL
#               )

```{r,model-interactions}
#| message: false
#| echo: false
#| label: fig-interaction-effects
#| warning: false
#| fig-cap: "Interaction Effects of models 1 and 2"
#| fig-width: 10
#| fig-height: 10

#cm<-c('Incomeunder_15000'='Under 15000')

d<-modelplot(model1,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant"))+
  scale_color_manual(values = c("black", "red"))+ggtitle("1st Model")+xlab("")

e<-modelplot(model2,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
  scale_color_manual(values = c("black", "red"))+ggtitle("2nd Model")

f<-modelplot(model3,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
  scale_color_manual(values = c("black", "red"))+ggtitle("3rd Model")

d+e+plot_layout(guides = 'collect',ncol=1)+plot_annotation(title="Interaction effects for models 1 and 2") & theme(legend.position = 'bottom')

```{r,third-model}
#| message: false
#| echo: false
#| label: fig-interaction-effects-mod3
#| warning: false
#| fig-cap: "Interaction Effects of model 3"
#| fig-width: 10
#| fig-height: 15

f+plot_layout(guides = 'collect',ncol=1)+plot_annotation(title="Interaction effects for model 3") & theme(legend.position = 'bottom')

**So far, considering that the interaction of socioeconomic factors are important predictors of vaccination status, I believe the model that provides a more cohesive picture of the data is model 3 as it shows differences between race and income which could be important if we consider the types of occupations that people in certain regions do, and how these influences their vaccination status.**

>>>>

#make variables of interest factors
clean_data<-clean_data %<>%
  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)

# recode the response of interest for categorical analysis
clean_data<-clean_data %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

#get health regions to eliminate
clean_data <-clean_data %>%
   filter(Health_Region!="North East" , Health_Region!="North West")

#eliminate the levels of the removed health regions
clean_data <- droplevels(clean_data)

## relevel for reference groups in the model

 clean_data<- within(clean_data, Age_group <- relevel(Age_group, ref = "16_24"))
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, Income <- relevel(Income, ref = "over_90000"))

@@

@@

# clean_data %>%
#   tbl_summary(
#     include=c(
#     Health_Region,)
#   ) %>% bold_labels()

@@

Health_Regions<-data.frame(Health_Region=c(
  #"North West",
  #"North East",
  "West",
  "East",
  "Central",
  "Toronto"),
  Freq=c(#232299,
    #557000,
    4095589,
    3742520,
    5032410,
    1440644
  )
)

@@

| Variable             | Values                                                                                                                          |
|:-----------------|------------------------------------------------------|
| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
|                      |                                                                                                                                 |

@@

The cleaning process also included removing outliers that were identified during the preliminary analyses. Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000. After cleaning, the dataset contained 5,247 entries (each entry corresponding to a unique respondent).

@@

For each survey participant certain data was automatically captured. This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset. Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset. Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN. Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset. Finally, we assigned the corresponding Health Region to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).

Following an assessment of the number of entries corresponding to each Health Region in the clean dataset, there were a relatively low number of observations from the North East and North West regions (2.9% and 1.4% of the total clean dataset, respectively). Due to the low number of entries in each case, we omitted the entries from these regions from further analyses. Therefore, the total number of unique entries used for analysis was 3,551.

@@

## Statistical model

Because of the binary outcome of the survey answer of our interest (vaccination status as "yes" or "no") we used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location. Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. At the same time in the case of Canada, others have indicated an ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].

Therefore, we built a logistic regression model that accounted for interactions between the available socio-economic factors obtained from the Fields Covid survey. The model appears in @eq-model3,

@@

Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, $\beta_1...\beta_7$ indicate the coefficients for each of the fixed and interaction effects of age group, race, Health Region, and income. The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.

@@

## the first model is a simple logistic regression with fixed effects only svyglm uses the glm
## function to calculate, so it does not allow for random effects.
# 
# model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
#               design=a1_rake,
#               family = quasibinomial(), 
#               control= list(maxit=25))
# 
# 
# model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
#               design=a1_rake,
#               family = quasibinomial(), 
#               control= list(maxit=25))

@@

#summary(model1.test)
#summary(model1.hr)
# 
# model_comp<-as_tibble(AIC(model1,model2,model3))%>%
#   select(-c(eff.p,deltabar))%>%
#   add_column(model=c("model1","model2","model3"))
# 
# model_comp%>%
#   kbl(booktabs=T,
#       col.names = c("AIC","model"))
#   
## helps visualize the model output

@@

data_summary %>%
  modify_header(label = "**Variable**") %>%
  as_kable_extra(booktabs = TRUE) %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down"` argument here.
  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)

@@

## Vaccination status, socio-economic and geographical variables

@tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors. The reference groups in each case were set as follow:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income). Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.58), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.17) and Other Race/Ethnicity (a group  that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000 and CAD 39,999.

Regarding the interaction of socio-economic and geographical factors, statistically significant odds ratios were the Central Health Region and Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older with a household income between 25000 and 39999 CAD (OR=0.001), and with an income between 40000 and 59999 (OR=3.12)

```{r,model-table}
#| message: false
#| echo: false
#| label: tbl-model
#| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"

# modelsummary(model3, shape=term~response)
# 
# 
# panels <- list(
#   "Health Regions" = list(
#     "(I)" = lm(mpg ~ 1, data = mtcars),
#     "(II)" = lm(mpg ~ qsec, data = mtcars)
#   ),
#   "Outcome: hp" = list(
#     "(I)" = lm(hp ~ 1, data = mtcars),
#     "(II)" = lm(hp ~ qsec, data = mtcars)
#   )
# )
# 
# modelsummary(
#   panels,
#   shape = "rbind",
#   gof_map = gm)

model3 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group ~ "Age Group")) %>%
    bold_labels()%>%
  as_kable_extra(booktabs = TRUE, longtable=T) %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down" argument here.
  kableExtra::kable_styling(latex_options = "repeat_header")
```
